---
title: "Supplementary Information for 'CoronaNet COVID-19 Government Response Event Dataset'"
header-includes:
   - \usepackage{lineno}
   - \linenumbers
csl: nature.csl
author: 
  - name: Cindy Cheng
  - name: Joan Barceló
  - name: Allison Spencer Hartnett
  - name: Robert Kubinec
  - name: Luca Messerschmidt
date: "May 16th, 2020"
toc: false
bibliography: BibTexDatabase.bib
always_allow_html: yes
output: 
  #bookdown::word_document2
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header:
          preamble.tex
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
urlcolor: blue
mask: yes
---

# Appendix A: Description of Dataset Fields {-}

The format of the data is in country-day-`record_id` format. Some `record_id` values have letters appended to indicate that the general policy category `type` also has a value for `type_sub_cat`, which contains more detail about the policy, such as whether health resources refers to masks, hospitals, or doctors. Some entries are marked as `new_entry` in the `entry_type` field for when a policy of that type was first implemented in the country. Later updates to those policies are marked as updates in `entry_type`. To see how policies are connected, look at the `policy_id` field for all policies from the first entry through each update for a given country/province/city. If an entry was corrected after initial data collection, it will read corrected in the `correct_type` field (the original incorrect data has already been replaced with the corrected data). 

1. **`coronanet_release.csv`** This file contains variables from the CoronaNet government response project, representing national and sub-national policy event data from more than 190 countries since December 31st, 2019. The data include a qualitiatve description for each recorded policy announcement, as well as information about different dimensions of the policy, including the type of policy, the policy initiator, the target of a policy, the policy targets (i.e. other countries), the level of enforcement and the enforcer of a policy.  For more detail on this data, see our [codebook here](https://docs.google.com/document/d/1zvNMpwj0onFvUZ_gLl4RRjqS-clbHv3TIX6EOHofsME).

2. **`coronanet_release_allvars.csv`** This file contains the government policy response information from `coronanet_release.csv` along with the following datasets:

    a. Tests from the CoronaNet testing database (See http://coronanet-project.org for more info);
    b. Cases/deaths/recovered from the [JHU data repository](https://github.com/CSSEGISandData/COVID-19);
    c. Country-level covariates including GDP, V-DEM democracy scores, human rights indices, power-sharing indices, and press freedom indices from the [Niehaus World Economics and Politics Dataverse](https://niehaus.princeton.edu/news/world-economics-and-politics-dataverse)
    
## `coronanet_release.csv` Field Dictionary {-}

1. `record_id` Unique identifier for each unique policy record
2. `policy_id` Identifier linking new policies with subsequent updates to policies
3. `recorded_date` When the record was entered into our data
4. `date_updated` When we can confirm the country - policy type was last checked/updated (we can only confirm  policy type for a given country is up to date as of this date)
5. `date_announced` When the policy is announced
6. `date_start` When the policy goes into effect
7. `date_end` When the policy ends (if it has an explicit end date)
8. `entry_type` Whether the record is new, meaning no restriction had been in place before, or an update (restriction was in place but changed). 
9. `correct_type` Whether a record as been corrected or not. 
10. `event_description` A short description of the policy change
11. `domestic_policy` Indicates where policy targets an area within the initiating country (i.e. is domestic in nature)
12. `type` The category of the policy
13. `type_sub_cat` The sub-category of the policy (if one exists)
14. `type_text` Any additional information about the policy type (such as the number of ventilators/days of quarantine/etc.)
15. `index_high_est` The high (95% posterior density) estimate of the country policy activity score (0-100)
16. `index_med_est` The median (most likely) estimate of the country policy activity score (0-100)
17. `index_low_est` The low (95% posterior density) estimate of the country policy activity score (0-100)
18. `index_country_rank` The relative rank by each day for each country on the policy activity score
19. `country` The country initiating the policy
20. `init_country_level` Whether the policy came from the national level or a sub-national unit
21. `province` Name of sub-national unit
22. `target_country` Which foreign country a policy is targeted at (i.e. travel policies)
23. `target_geog_level` Whether the target of the policy is a country as a whole or a sub-national unit of that country
24. `target_region` The name of a regional grouping (like ASEAN) that is a target of the policy (if any)
25. `target_province` The name of a province targeted by the policy (if any)
26. `target_city` The name of a city targeted by the policy (if any)
27. `target_other` Any geographical entity that does not fit into the targeted categories mentioned above
28. `target_who_what` Who the policy is targeted at
29. `target_direction` Whether a travel-related policy affects people coming in (Inbound) or leaving (Outbound)
30. `travel_mechanism` If a travel policy, what kind of transportation it affects
31. `compliance` Whether the policy is voluntary or mandatory
32. `enforcer` What unit in the country is responsible for enforcement
33. `link` A link to at least one source for the policy
34. `ISO_A3` 3-digit ISO country codes
35. `ISO_A2` 2-digit ISO country codes

## `coronanet_release_allvars.csv` Field Dictionary {-}

1. All of the fields listed above, plus
2. `tests_daily_or_total` Whether a country reports the daily count of tests a cumulative total
3. `tests_raw` The number of reported tests collected from host country websites or media reports
4. `deaths` The number of COVID-19 deaths, aggregated to the country-day level (JHU CSSE data)
5. `confirmed_cases` The number of confirmed cases of COVID-19, aggregated to the country-day level (JHU CSSE data)
6. `recovered` The number of recoveries from COVID-19,  aggregated to the country-day level (JHU CSSE data)
7. `ccode` The Correlates of War country code
8. `ifs` IMF IFS country code

9. `Rank_FP` (most recent year available from Niehaus dataset) Reporters without Borders Press Freedom Annual Ranking
10. `Score_FP` (most recent year available from Niehaus dataset) Reporters with Borders Press Freedom Score
11. `state_IDC` (most recent year available from Niehaus dataset) State/Provincial Governments Locally Elected
12. `muni_IDC` (most recent year available from Niehaus dataset) Municipal Governments Locally Elected
13. `dispersive_IDC` (most recent year available from Niehaus dataset) Dispersive Powersharing 
14. `constraining_IDC` (most recent year available from Niehaus dataset) Constraining Powersharing 
15. `inclusive_IDC` (most recent year available from Niehaus dataset) Inclusive powersharing 
16. `sfi_SFI` (most recent year available from Niehaus dataset) State fragility index
17. `ti_cpi_TI` (most recent year available from Niehaus dataset) Corruption perceptions index
18. `pop_WDI_PW` (most recent year available from Niehaus dataset) World Bank population
19. `gdp_WDI_PW` (most recent year available from Niehaus dataset) World Bank GDP (total)
20. `gdppc_WDI_PW` (most recent year available from Niehaus dataset) World Bank GDP per capita
21. `growth_WDI_PW` (most recent year available from Niehaus dataset) World Bank GDP growth percent
22. `lnpop_WDI_PW` (most recent year available from Niehaus dataset) Log of World Bank population
23. `lngdp_WDI_PW` (most recent year available from Niehaus dataset) Log of World Bank GDP
24. `lngdppc_WDI_PW` (most recent year available from Niehaus dataset) Log of World Bank GDP per capita
25. `disap_FA` (most recent year available from Niehaus dataset) 3 category, ordered variable for disappearances index
26. `polpris_FA` (most recent year available from Niehaus dataset) 3 category, ordered variable for political imprisonment index
27. `latentmean_FA` (most recent year available from Niehaus dataset) the posterior mean of the latent variable index for human rights protection)
28. `transparencyindex_HR` (most recent year available from Niehaus dataset) Transparency Index
29. `EmigrantStock_EMS` (most recent year available from Niehaus dataset) Total emigrant stock from
30. `v2x_polyarchy_VDEM` (most recent year available from Niehaus dataset) Electoral democracy index
31. `news_WB` (most recent year available from Niehaus dataset) Daily newspapers (per 1,000 people)


# Appendix B: Research Assistant Training and Management {-}


## RA Training {-}


In order to register for this data collection effort, RAs watch a mandatory 2 hour video training of the survey instrument which explains how to use the survey to document new policies as well as to update, and correct previous entries. 

RAs are also provided with written guidelines on how to collect data and a comprehensive codebook. To briefly describe it here, the written guidelines provide a definition of what counts as a new or updated policy (see the Data Schema section for more details) and provides a checklist for RAs to follow in order to identify and document different policies. In the checklist, RAs are instructed to find policies by checking the sources in the order given in the guidelines to identify policies, to document the relevant information into the survey and to save and upload a document of the source they found for each policy into the Qualtrics survey. Meanwhile, the codebook provides descriptions and examples of the different possible response options in the survey. 

Using a training video and the written codebook also has the added benefit of helping us efficiently disseminate the information RAs need to use the survey experiment consistently. Before getting allocated to a country based on language skills and interests, RAs participate in an online test where they are asked a randomized set of questions about how to code different policies. To pass the test, RAs need to get at least 70% of the entries correct.

Overall, in order to participate as an RA in this project, RAs must fill out a consent form ^[See here for the [link to the form](https://docs.google.com/forms/d/e/1FAIpQLSeybAW0DC0UE1x2EqLiTifVFuSUxqJLGFB8VI4wVCG61tVYKg/viewform).] in which:

* They identify themselves.
* They certify that they have viewed the training video and passed the test.
* They certify they have joined the CoronaNet Slack Channel (see section below for more information).
* They certify that they understand that RA responsibilities entail 
  + gathering historical data on COVID-19 government policy actions for their country, and;
  + providing daily updates for new government policy actions.
* They certify that they understand they can access the data collection guidelines and codebook or pose their questions on the Slack Channel.
* They certify that they are expected to upload .pdfs of the sources they access to the survey instrument.

Once the RA submits the form, they are sent a personalized link to access the survey. With the customized link, we are also able to keep track of which RA coded what entries and when the last policy entry was.

## Real-Time Communication and Feedback {-}

Once an RA joins the project, they can pose their questions on a CoronaNet Slack channel, which they must join in order to participate in the project. The channel allows any RA to pose a question or raise an issue they may have in using the survey instrument to any of the PIs and allows all other RAs to learn from the exchange at the same time. As such, RAs are able to receive feedback and learn from each other's questions in a timely and centralized manner. Since the data collection effort was launched on March 28, 2020 both RAs and PIs have actively used Slack to communicate with one another.  On the Slack channel devoted to asking questions about the Qualtrics data survey in particular, there were more 2600 messages posted by 154 project members at the time of writing. 

<!---To provide a better sense of the level of activity on Slack, Figure \@ref(fig:members) plots the number of project members who had logged into the Slack per day as well as the number of project members who posted at least one message per day and shows that participation and activity Slack  has steadily grown over time.^[Note, the dip in the overall trend corresponds to weekend days.]

--->


# Appendix C: List of Contributors to Dataset {-}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message=FALSE,tinytex.verbose = TRUE)
library(readr)
library(knitr)
library(magrittr)
library(dplyr)
require(kableExtra)
```

```{r ratable}


contribution <- read_csv("../data/CoronaNet/People/contribution_clean.csv")


contribution <- dplyr::rename(contribution, Country = "country") %>% 
  mutate(Vita=coalesce(Vita,""),
         Affiliation=coalesce(Affiliation,"")) 
knitr::kable(contribution,booktabs=T,longtable=T,
          caption = 'Contributing Researchers and their Responsible Countries') %>% 
  kable_styling(latex_options = c("striped", "hold_position")) %>% 
  column_spec(2:3,width="2cm") %>% 
  column_spec(4,width="3cm")


```
